---
title: "RIMERMAN_HW10"
author: "Mitch Rimerman"
date: "2022-11-15"
output: pdf_document
---

```{r setup, include=FALSE, results=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
```

## Problem 1

Consider the “Rosenbrock” function.
\[
f(x)_1,x_2) = 100(x_2-x^2_1)^2 + (1-x_1)^2.
\]
With a starting point $[0,0]^T$, apply two iterations of Newton’s method to minimize Rosenbrock
function.

\[
x^{(k+1)}:=x^{(k)}-F(x^{(k)})^{-1}g^{(k)}
\]
Where
\[
g^{(k)} = \nabla f(x_1,x_2)=\begin{bmatrix}
400x_1^3-400x_1x_2+2x_1-2 \\
200x_2-200x_1^2
\end{bmatrix}\quad\text{ at }x^{(k)}
\]
and
\[
F(x^{(k)})=\nabla^2f(x_1,x_2)=\begin{bmatrix}
1200x_1^2-400x_2+2 & -400x_1 \\
-400x_1 &  200
\end{bmatrix}\quad\text{ at }x^{(k)}
\]

### First Iteration
\[
g^{(0)} = \nabla f(0,0)=\begin{bmatrix}
-2 \\
0
\end{bmatrix}\quad\text{ and }\quad F(x^{(0)})=\nabla^2f(0,0)=\begin{bmatrix}
2 & 0 \\
0 & 200
\end{bmatrix}
\]
\[
F(x^{(0)})^{-1}=\frac{1}{400}\begin{bmatrix}
200 & 0 \\
0 & 2
\end{bmatrix}
\]
\[
x^{(1)}=\begin{bmatrix}
0\\
0
\end{bmatrix}-\begin{bmatrix}
1/2 & 0 \\
0 & 1/200
\end{bmatrix}\begin{bmatrix}
-2 \\
0
\end{bmatrix}=\begin{bmatrix}
0\\
0
\end{bmatrix}-\begin{bmatrix}
-1\\
0
\end{bmatrix}=\begin{bmatrix}
1\\
0
\end{bmatrix}
\]

### Second Iteration
\[
g^{(1)} = \nabla f(1,0)=\begin{bmatrix}
400 \\
-200
\end{bmatrix}\quad\text{ and }\quad F(x^{(1)})=\nabla^2f(1,0)=\begin{bmatrix}
1202 & -400 \\
-400 &  200
\end{bmatrix}
\]
\[
F(x^{(1)})^{-1}=\frac{1}{80400}\begin{bmatrix}
200 & 400 \\
400 & 1202
\end{bmatrix}
\]
\[
x^{(2)}=\begin{bmatrix}
1\\
0
\end{bmatrix}-\frac{1}{80400}\begin{bmatrix}
200 & 400 \\
400 & 1202
\end{bmatrix}\begin{bmatrix}
400 \\
-200
\end{bmatrix}=\begin{bmatrix}
1\\
0
\end{bmatrix}-\frac{1}{80400}\begin{bmatrix}
0\\
-80400
\end{bmatrix}=\begin{bmatrix}
1\\
0
\end{bmatrix}-\begin{bmatrix}
0\\
-1
\end{bmatrix}=\begin{bmatrix}
1\\
1
\end{bmatrix}
\]

Thus, the first two iterates are $[1,0]^T$, followed by $[1,1]^T$.

\newpage
## Problem 2
Let $S=\text{span}\{x_1,x_2,x_3\}$, where
\[
x_1=\begin{bmatrix}
2\\
1\\
0\\
0
\end{bmatrix},\quad x_2=\begin{bmatrix}
2\\
2\\
0\\
-3
\end{bmatrix},\quad x_3=\begin{bmatrix}
0\\
-1\\
1\\
0
\end{bmatrix}
\]
Find an orthonormal basis for $S$, using the Gram-Schmidt algorithm
\[
\tilde{q}_1=x_1\neq0\quad\Rightarrow\quad q_1=\frac{x_1}{\lVert x_1\rVert}=\frac{1}{\sqrt{5}}\begin{bmatrix}
2\\
1\\
0\\
0
\end{bmatrix}=\begin{bmatrix}
2/\sqrt{5}\\
1/\sqrt{5}\\
0\\
0
\end{bmatrix}
\]
\begin{align*}
\tilde{q}_2 & =x_2-(q_1^Tx_2)q_1 \\
&=\begin{bmatrix}
2\\
2\\
0\\
-3
\end{bmatrix}-\left(\begin{bmatrix}
2/\sqrt{5} & 1/\sqrt{5} & 0 & 0
\end{bmatrix}\begin{bmatrix}
2\\
2\\
0\\
-3
\end{bmatrix}\right)\begin{bmatrix}
2/\sqrt{5}\\
1/\sqrt{5}\\
0\\
0
\end{bmatrix} \\
&= \begin{bmatrix}
2\\
2\\
0\\
-3
\end{bmatrix}-\frac{6}{\sqrt{5}}\begin{bmatrix}
2/\sqrt{5}\\
1/\sqrt{5}\\
0\\
0
\end{bmatrix} \\
&= \begin{bmatrix}
2\\
2\\
0\\
-3
\end{bmatrix}-\begin{bmatrix}
12/5 \\
6/5 \\
0\\
0
\end{bmatrix} \\
&= \begin{bmatrix}
-2/5 \\
4/5 \\
0 \\
-3
\end{bmatrix}\neq 0
\end{align*}
\[
q_2=\frac{\tilde{q}_2}{\lVert \tilde{q}_2\rVert}=\frac{\sqrt{5}}{7}\begin{bmatrix}
-2/5 \\
4/5 \\
0 \\
-3
\end{bmatrix}=\begin{bmatrix}
-2\sqrt{5}/35\\
4\sqrt{5}/35\\
0\\
-3\sqrt{5}/7
\end{bmatrix}
\]
\begin{align*}
\tilde{q}_3 &=x_3-(q_1^Tx_3)q_1-(q_2^Tx_3)q_2 \\
&=\begin{bmatrix}
0\\
-1\\
1\\
0
\end{bmatrix}-\left(\begin{bmatrix}
2/\sqrt{5} \\
1/\sqrt{5} \\
0 \\ 
0
\end{bmatrix}^T\begin{bmatrix}
0\\
-1\\
1\\
0
\end{bmatrix}\right)\begin{bmatrix}
2/\sqrt{5}\\
1/\sqrt{5}\\
0\\
0
\end{bmatrix}-\left(\begin{bmatrix}
-2\sqrt{5}/35\\
4\sqrt{5}/35\\
0\\
-3\sqrt{5}/7
\end{bmatrix}^T\begin{bmatrix}
0\\
-1\\
1\\
0
\end{bmatrix}\right)\begin{bmatrix}
-2\sqrt{5}/35\\
4\sqrt{5}/35\\
0\\
-3\sqrt{5}/7
\end{bmatrix} \\
&=\begin{bmatrix}
0\\
-1\\
1\\
0
\end{bmatrix}-\frac{-1}{\sqrt{5}}\begin{bmatrix}
2/\sqrt{5}\\
1/\sqrt{5}\\
0\\
0
\end{bmatrix}-\frac{-4\sqrt{5}}{35}\begin{bmatrix}
-2\sqrt{5}/35\\
4\sqrt{5}/35\\
0\\
-3\sqrt{5}/7
\end{bmatrix} \\
&=\begin{bmatrix}
0\\
-1\\
1\\
0
\end{bmatrix}-\begin{bmatrix}
-2/5\\
-1/5\\
0\\
0
\end{bmatrix}-\begin{bmatrix}
8/245\\
-16/245\\
0\\
12/49
\end{bmatrix} \\
&=
\begin{bmatrix}
18/49 \\
13/49 \\
1 \\
-12/49
\end{bmatrix}
\end{align*}
\[
q_3=\frac{\tilde{q}_3}{\lVert \tilde{q}_3\rVert}=\frac{7\sqrt{85}}{85}\begin{bmatrix}
18/49 \\
-36/49 \\
1 \\
-12/49
\end{bmatrix}=\begin{bmatrix}
18\sqrt{85}/595\\
-36\sqrt{85}/595\\
7\sqrt{85}/85\\
-12\sqrt{85}/595
\end{bmatrix}
\]
Thus, the set of orthonormal vectors is
\[
\left\{\begin{bmatrix}
2/\sqrt{5}\\
1/\sqrt{5}\\
0\\
0
\end{bmatrix},\begin{bmatrix}
-2\sqrt{5}/35\\
4\sqrt{5}/35\\
0\\
-3\sqrt{5}/7
\end{bmatrix},\begin{bmatrix}
18\sqrt{85}/595\\
-36\sqrt{85}/595\\
7\sqrt{85}/85\\
-12\sqrt{85}/595
\end{bmatrix} \right\}
\]

\newpage
## Problem 3
Write a code and implement the Gauss-Newton Method on the last example given in the
lecture, to find $A$, $\omega$, and $\phi$ such that the resulting sinusoid
\[y = A\text{ sin}(\omega t + \phi)\]
best fits $(t_i,y_i)$,$i = 1,2,...,21$, with $t_1 = 0$ and $t_{21} = 10$ and $y_i$ given roughly below.

Suppose the points in the graph are as follows

```{r p3a, echo=FALSE}
data <- data.frame((1:21)/2-.5,c(1.15,1.85,1.95,1.9,1.3,.1,-.5,-1.45,-1.95,-2.1,-1.4,-.35,.35,1.4,1.85,2.05,1.7,.75,-.2,-1.05,-1.6))
names(data) <- c("t","y")
data
ggplot(data)+
  geom_point(aes(x=t,y=y), color="#2D4790")+
  labs(x="Time",y="Measurement")+
  theme_minimal()
```

```{r p3b}
results <- data.matrix(data.frame(1,1,1))
names(results) <- c("A","omega","phi")
jacobian <- function(x){
  a <- -sin(x[2]*data["t"]+x[3])
  b <- -x[1]*data["t"]*cos(x[2]*data["t"]+x[3])
  c <- -x[1]*cos(x[2]*data["t"]+x[3])
  data.matrix(data.frame(a,b,c))
}

normdiff = 1
i = 1
while (normdiff > 10^(-50)){
  r <- data.matrix(data.frame(data["y"]-results[i,1]*sin(results[i,2]*data["t"]+results[i,3])))
  results <- rbind(results, t(data.matrix(results[i,])-solve(t(jacobian(results[i,])) %*% jacobian(results[i,])) %*% t(jacobian(results[i,])) %*% r))
  normdiff <- norm(results[i+1],type="2")-norm(results[i],type="2")
  i <- i+1
}

estimation <- data.frame((1:2001)/200-.005)
names(estimation) <- "t"
estimation["b"] <- results[7,1]*sin(results[7,2]*estimation["t"]+results[7,3])

results[7,]
ggplot()+
  geom_point(data=data, aes(x=t,y=y), color="#2D4790")+
  geom_line(data=estimation, aes(x=t,y=b), color="#2D904E")+
  labs(title="Gauss-Newton Estimation",x="Time",y="Measurement")+
  theme_minimal()

```


