---
title: "Rimerman_HW12"
author: "Mitch Rimerman"
date: "2022-12-02"
output: pdf_document
---

```{r setup, include=FALSE, results=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
```

## Problem 1
### a

\[
f(x)=\sin x-\cos x
\]

\[
f'(0)=\cos(0)+\sin(0)=1+0=1
\]

\[
f''(x)=\cos x-\sin x\quad\Rightarrow\quad \max\{|f''(x)|: x\in\mathbb{R}\}=\sqrt{2}=M
\]

```{r p1a}
h <- 10^(-(1:12))
f <- function(x){
  sin(x)-cos(x)
}

Fh <- rep(0,12)
error <- rep(0,12)
Mh <- rep(0,12)
i=12
for (i in 1:12) {
  Fh[i] <- (f(0+h[i])-f(0))/h[i]
  error[i] <- 1-Fh[i]
  Mh[i] <- sqrt(2)*h[i]/2
}
answertable <- data.frame(h,Fh,error,Mh)
names(answertable) <- c("h","Fh(0)","f'(0)-Fh(0)","Mh/2")
knitr::kable(answertable, "simple", scientific=TRUE, digits=15)
```

The minimum error corresponds to the theoretical error as it is less than $Mh/2$ until $h<10^{-8}$, where rounding errors make the true error higher than the max expectation, which is especially prevalent when $h=10^{-12}$ (on my computer).

### b

\[
f(x)=(1+x)^{-1}
\]
\[
f'(1)=-(1+1)^{-2}=-1/4
\]
\[
f'''(x)=-6(1+x)^{-4}\quad\Rightarrow\quad \max\{|f'''(x)|:x\in[0,2]\}=6=M
\]

```{r p1b}
h <- 10^(-(1:12))
f <- function(x){
  1/(1+x)
}

Fh <- rep(0,12)
error <- rep(0,12)
Mh <- rep(0,12)
i=12
for (i in 1:12) {
  Fh[i] <- (f(1+h[i])-f(1-h[i]))/(2*h[i])
  error[i] <- (-1/4)-Fh[i]
  Mh[i] <- h[i]^2
}
answertable <- data.frame(h,Fh,error,Mh)
names(answertable) <- c("h","Fh(0)","f'(0)-Fh(0)","h^2*M/6")
knitr::kable(answertable, "simple", scientific=TRUE, digits=30)
```

Similar to part a, the minimum error corresponds to the theoretical error as it is less than $h^2M/6$ until $h<10^{-6}$, where rounding errors make the true error higher than the max expectation, which is especially prevalent when $h=10^{-12}$ (on my computer).

Note in this case $F_h(x)=\frac{f(x+h)-f(x-h)}{2h}$.

\newpage
## Problem 2
\[
\int x^2 \ln(x)dx= \frac{x^3\ln(x)}{3}-\int\frac{x^2}{3}=\frac{x^3\ln(x)}{3}-\frac{x^3}{9}=\frac{x^3}{9}(3\ln(x)-1)
\]
\[
\int_1^3 x^2 \ln(x)dx=\left[\frac{x^3}{9}(3\ln(x)-1)\right]_1^3=9\ln(3)-\frac{26}{9}\approx6.9986217
\]

### a

```{r p2a}
f <- function(x){
  x^2*log(x)
}

N <- 16
h <- 2/N
a <- 1+((1:N)-1)*h
b <- 1+(1:N)*h
approxint <- data.frame((h/2)*(sum(f(a))+sum(f(b))),abs((h/2)*(sum(f(a))+sum(f(b)))-(9*log(3)-26/9)))
names(approxint) <- c("Approximate Intergral","Error")
knitr::kable(approxint, "simple", scientific=TRUE, digits=30)

N <- 32
h <- 2/N
a <- 1+((1:N)-1)*h
b <- 1+(1:N)*h
approxint <- data.frame((h/2)*(sum(f(a))+sum(f(b))),abs((h/2)*(sum(f(a))+sum(f(b)))-(9*log(3)-26/9)))
names(approxint) <- c("Approximate Intergral","Error")
knitr::kable(approxint, "simple", scientific=TRUE, digits=30)

```

### b

```{r p2b}
f <- function(x){
  x^2*log(x)
}

N <- 16
h <- 2/N
a <- 1+((1:N)-1)*h
b <- 1+(1:N)*h
c <- (a+b)/2
simpson <- f(a)+4*f(c)+f(b)
approxint <- sum(simpson)*h/6
approxint <- data.frame(sum(simpson)*h/6,abs(sum(simpson)*h/6-(9*log(3)-26/9)))
names(approxint) <- c("Approximate Intergral","Error")
knitr::kable(approxint, "simple", scientific=TRUE, digits=30)

N <- 32
h <- 2/N
a <- 1+((1:N)-1)*h
b <- 1+(1:N)*h
c <- (a+b)/2
simpson <- f(a)+4*f(c)+f(b)
approxint <- sum(simpson)*h/6
approxint <- data.frame(sum(simpson)*h/6,abs(sum(simpson)*h/6-(9*log(3)-26/9)))
names(approxint) <- c("Approximate Intergral","Error")
knitr::kable(approxint, "simple", scientific=TRUE, digits=30)
```

