---
title: "GE03"
author: "Mitch Rimerman"
date: "10/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Import

We must first load the necessary packages and import the data sets.

```{r data}
library(tidyverse)
library(viridis)
all_states <- map_data("state")
congress <-read_csv("womenincongress.csv")
electionData <- read_csv("2012.csv")
```

## Graph of Women Representatives

### Data Manipulation

To combine the state map data with the representative data, we must first rename the state column and then use the join command.

```{r manip1}
names(congress)[2] <- "region"
stateData <- left_join(all_states,congress,by="region")
```

Now, we must create a new variable for the proportion of women representatives per state.

```{r repProp}
stateData <- mutate(stateData, repProp=representatives/total)
```

### Graph

With this new variable, we can create a graph of the proportion of women representatives:

```{r graph1}
ggplot()+
  geom_polygon(data=stateData,aes(x=long, y=lat, group = group, fill=repProp),color="grey50")+
  coord_map()+
  labs(x="",y="",title="Women in the House")+
  scale_fill_gradient(name="Female Representatives",high="darkred",low="white") +
  theme_classic()+ 
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(), axis.ticks.x =element_blank(), 
        axis.text.x = element_blank())
```

## Graph of 2012 Election in the South

### Data Manipulation

Again, we must add a few variables and combine the data sets.

```{r manip2}
names(electionData)[1] <- "region"

electionData$ObamaPerc <- electionData$ObamaVotes/(electionData$ObamaVotes+electionData$RomneyVotes+electionData$JohnsonVotes+electionData$SteinVotes)
electionData$RomneyPerc <- electionData$RomneyVotes/(electionData$ObamaVotes+electionData$RomneyVotes+electionData$JohnsonVotes+electionData$SteinVotes)
electionData <- merge(all_states,electionData,by="region")
```

Now, we must define the South region, and create a new data set that only includes those states.

```{r south}
South <- filter(electionData, region %in% c("texas", "oklahoma", "louisiana", "arkansas", "alabama", "mississippi", "tennessee", "kentucky", "west virginia", "delaware", "maryland", "district of columbia", "virginia", "north carolina", "south carolina", "georgia", "florida"))
```

Now, we can create our graph of the 2012 election in the south:

```{r graph2}
ggplot()+
  geom_polygon(data=South,aes(x=long, y=lat, group = group, fill=ObamaPerc),color="grey50")+
  coord_map()+
  labs(x="",y="",title="2012 Election Results")+
  theme_classic()+ 
  theme(axis.ticks.y = element_blank(),axis.text.y = element_blank(), axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
  scale_fill_gradient2(name="Obama's Percenatage",low="red",mid="white",high="blue",midpoint=.5)
```




